---
title: "A3_QMM_Transportation_model"
author: "Nikhila Reddy Vantari"
date: "2024-10-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question:

Heart Start produces automated external defibrillators (AEDs) in each of two different plants (A
and B). The unit production costs and monthly production capacity of the two plants are
indicated in the table below. The AEDs are sold through three wholesalers. The shipping cost
from each plant to the warehouse of each wholesaler along with the monthly demand from each
wholesaler are also indicated in the table. How many AEDs should be produced in each plant,
and how should they be distributed to each of the three wholesaler warehouses so as to minimize
the combined cost of production and shipping?


```{r}
# Load packages - lp solve
library(lpSolve)
```


```{r}
#Formatting a data frame for all capacities and production costs
t.data <- data.frame(Warehouse_1 = c("$22", "$16", 80),
                     Warehouse_2 = c("$14", "$20", 60),
                     Warehouse_3 = c("$30", "$24", 70),
                     Unit_production_cost = c("$600", "$625", " "),
                     Monthly_production_capacity = c(100, 120, " "))
rownames(t.data) <- c("Plant A", "Plant B", "Monthly Demand")


library(knitr)
library(kableExtra)

t.data %>%
  kable() %>%
  kable_classic() %>%
  column_spec(1, border_right = TRUE) %>%
  column_spec(4, border_right = TRUE) %>%
  row_spec(2, extra_css = "border-bottom:dotted") %>%
  add_header_above(header = c(" " = 1, "Unit Shipping Cost" = 3, " " = 2))
```



## Decision Variables:

aij as the quantity transported from Plant A to Warehouse j (j = 1, 2, 3)
bij as the quantity transported from Plant B to Warehouse j (j = 1, 2, 3)


## Objective Function:
Our goal is to minimize total expenses, including both manufacturing and transportation costs:

Minimize T = (600 + 22)a1 + (600 + 14)a2 + (600 + 30)a3 +
(625 + 16)b1 + (625 + 20)b2 + (625 + 24)b3
This can be simplified to:
Minimize T = 622a1 + 614a2 + 630a3 + 641b1 + 645b2 + 649b3

## Constraints:
### Supply Limitations:


Plant A: a1 + a2 + a3 ≤ 100
Plant B: b1 + b2 + b3 ≤ 120

### Demand Requirements:

Warehouse 1: a1 + b1 = 80
Warehouse 2: a2 + b2 = 60
Warehouse 3: a3 + b3 = 70

### Non-Negative Quantities:
a1, a2, a3, b1, b2, b3 ≥ 0



```{r}
#Formulating the shipping costs
costs_for_shipping <- c(600 + 22, 600 + 14, 600 + 30, 
           625 + 16, 625 + 20, 625 + 24)
```


```{r}
#Constraints in matrix
constraints <- matrix(c(1, 1, 1, 0, 0, 0,   # Capacity for Factory 1
                              0, 0, 0, 1, 1, 1,   # Capacity for Factory 2
                              1, 0, 0, 1, 0, 0,   # Demand for Warehouse A
                              0, 1, 0, 0, 1, 0,   # Demand for Warehouse B
                              0, 0, 1, 0, 0, 1),  # Demand for Warehouse C
                            nrow = 5, byrow = TRUE)
```


```{r}
#Establishing RHS values
rhs_values <- c(100, 120, 80, 60, 70)
```


```{r}
# Define the direction of each constraint (<= for capacities, >= for demands)
constraint_direction <- c("<=", "<=", ">=", ">=", ">=")

```


```{r}
# Solve the transportation problem using lp()
result <- lp("min", costs_for_shipping, constraints, constraint_direction, rhs_values)
```


```{r}
if (result$status == 0) {
  # If an optimal solution is found, display the results
  cat("Optimal solution found:\n")
  result$solution_matrix <- matrix(result$solution, nrow = 2, byrow = TRUE)
  rownames(result$solution_matrix) <- c("Factory 1", "Factory 2")
  colnames(result$solution_matrix) <- c("Warehouse A", "Warehouse B", "Warehouse C")
  print(result$solution_matrix)
}
```
```{r}
# Showcasing the output results for individual plants
total_plant_A <- sum(result$solution_matrix[1, ])
print(paste("Number of AEDs to be produced by plant 1: ", total_plant_A))

total_plant_B <- sum(result$solution_matrix[2, ])
print(paste("Number of AEDs to be produced by plant 2: ", total_plant_B))
```


```{r}
# Print the total minimized cost
print(paste("Minimum total cost:", result$objval))
```

## Conclusion:

Based on the transportation model optimization, the most cost-effective solution for Heart Start's AED production and distribution is as follows: Plant A should produce 100 AEDs, operating at full capacity, with 60 units shipped to Warehouse B and 40 units to Warehouse C. Plant B should produce 110 AEDs, slightly under its maximum capacity of 120, distributing 80 units to Warehouse A and 30 units to Warehouse C. This distribution strategy satisfies all demand requirements: 80 units for Warehouse A, 60 for Warehouse B, and 70 for Warehouse C. By implementing this optimal plan, Heart Start can minimize its combined production and shipping costs to $132,790, ensuring efficient resource utilization and meeting all wholesaler demands.
